---
title: "Artic Cod - Ratliff"
output: html_document
---


 explore factors (including year, location, bottom depth, bottom temperature, surface temperature) that are related to species abundance (measured by kg/hectare in the variable WTCPUE and #/hectare in the variable NUMCPUE). A particular interest is in exploring whether there are any trends in time, temperature, or other factors in the data. 

1. Introduction and purpose

2. Data cleaning

```{r, eval=FALSE}

##TODO: load in every data set even with giant files and subset them and then combine, add indicator variable for which place they were discovered in 

EBS82_84 <- read.csv("easternberingseashelf1982_1984.csv")
EBS85_89 <- read.csv("easternberingseashelf1985_1989.csv")
EBS90_94 <- read.csv("easternberingseashelf1990_1994.csv")
EBS95_99 <- read.csv("easternberingseashelf1995_1999.csv")
EBS00_04<- read.csv("easternberingseashelf2000_2004.csv")
EBS05_08 <- read.csv("easternberingseashelf2005_2008.csv")
EBS09_12 <- read.csv("easternberingseashelf2009_2012.csv")
EBS13_16 <- read.csv("easternberingseashelf2013_2016.csv")
EBS17 <- read.csv("easternberingseashelf2017.csv")
NBS82_17 <- read.csv("northberingseashelf1982_2017.csv")

mega <- rbind(EBS82_84, EBS85_89, EBS90_94, EBS95_99, EBS00_04, EBS05_08, EBS09_12, EBS13_16, EBS17, NBS82_17)
less <- mega[mega$COMMON == "Arctic cod" | mega$SCIENTIFIC == "Boreogadus saida",]
```


```{r}


#difficult to perform mean imputation with categorical items (stratum), seems to be MCAR no trends so just ignore it
data <- read.csv("cod.csv")
data[data==-9999.0] <- NA
data <- data[complete.cases(data)==TRUE,]
data$STATION = as.factor(data$STATION)
data$STRATUM = as.factor(data$STRATUM)
```


3. The model

```{r}
mod <- lm(WTCPUE ~ LATITUDE +LONGITUDE+ STATION+ STRATUM+ YEAR+ BOT_DEPTH+  BOT_TEMP+ SURF_TEMP+ VESSEL+ CRUISE+ HAUL, data = data)
step <- step(mod)
```

